Event History Analysis for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-12-11 manuscript/manuscript.Rmd
update ms.

# R references
r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation
my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE)

```

>>>>

# R references
r_refs(file = "r-references.bib", append=F) # append=F prevents multiple copies of the same citation
my_r_citation <- cite_r(file = "r-references.bib", footnote = TRUE) 
```

### 2024-12-11 manuscript/manuscript.tex
update ms.

\abstract{%
Time-to-event data such as response times, saccade latencies, and fixation durations form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition. However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration. The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data. Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states. One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches. To help achieve broader uptake, in this paper we outline a set of tutorials that detail how to implement one distributional method known as discrete-time EHA. We illustrate how to wrangle raw data files and calculate descriptive statistics, as well as how to calculate inferential statistics via Bayesian and frequentist multilevel regression modelling. Along the way, we touch upon several key aspects of the workflow, such as how to specify regression models, the implications for experimental design, as well as how to manage inter-individual differences. We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work. Finally, the project is written in R and freely available, which means the general approach can easily be adapted to other data sets, and all of the tutorials are available as .html files to widen access beyond R-users.
}

@@

In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a. survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003). We hope to show the value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience. Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.

@@

We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project. The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data sets. Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes. Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R. Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.

@@

For a comprehensive background context to EHA, we recommend several excellent textbooks (Allison, 2010; Singer \& Willett, 2003). Likewise, for a general introduction to understanding regression equations, we recommend several excellent textbooks (Gelman, Hill, \& Vehtari, 2020; Winter, 2019). Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before. Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology. As such, we only supply regression equations in section D of the supplementary material.

@@

Figure 3 shows that the early upswing in hazard is equal for both priming conditions, and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition. These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed). During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition. For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.

@@

We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
Functional programming basically means that you avoid writing your own loops but instead use functions that have been built and tested by others.
In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of descriptive EHA a lot quicker and more efficient.

@@

Our lab adopts a estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020). We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013). This means that wherever possible we include varying intercepts and slopes per participant
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020). We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

@@

In contrast, the `person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial. Therefore, the `person-trial-bin' data set generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin. This data set is used when fitting hazard models (Tutorials 2a and 3a). It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest. Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.

@@

\caption{Estimated discrete-time hazard, survivor, probability mass, and conditional accuracy functions for participant 6. Vertical dotted lines indicate the estimated median RTs. Error bars represent +/- 1 standard error of the respective proportion.}\label{fig:eha-plot}
\end{figure}

@@

These distributional results suggest that the participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus. Qualitatively similar results were obtained for the other five participants. When participants show qualitatively the same distributional patterns, one might consider to aggregate their data and make one plot (see Tutorial\_1a.Rmd).

@@

In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data set used in Tutorial 1a. Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).

@@

There are several analytic decisions one has to make when fitting a discrete-time hazard model. First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant. Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material). The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003). Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor. And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding. With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level. Brm() will then construct k-1 indicator variables (see model M1d in Tutorial\_2a.Rmd for an example). With index coding, one constructs an index variable which contains integers that correspond to different categories (see models M0i and M1i below). As explained by McElreath (2020), the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.

@@

If we want to study the average effect of prime type on hazard, uncontaminated by inter-individual differences, we can base our conclusion on Figure 7A. The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}). For the other bins, the 95\% credible interval contained zero.
The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin (320,360{]} (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin (360,400{]} (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin (400,440{]} (95\% CrI = {[}-0.40, -0.03{]}). For the other bins, the 95\% credible interval contained zero.

@@

In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data set used in Tutorial 1a. The general process is similar to Tutorial 2a, except that (a) we use the person-trial data set, (b) we use the logit link function, and (c) we change the priors. To keep the tutorial short, we only fitted the effects of model M1i (see Tutorial 2a) in the conditional accuracy model called M1i\_ca.

@@

In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data set used in Tutorial 1a. The general process is similar to that in Tutorial 2a, except that there are no priors to set.

To keep this tutorial short, we only fitted the effects from model M1i (see Tutorial 2a) using the function glmer() from the R package lme4. Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024). The resulting hazard model is called M1i\_f.

In Figure 11 we compare the parameter estimates of model M1i from brm() with those of model M1i\_f from glmer().

@@

Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure. However, model M1i\_f did not converge and resulted in a singular fit. This is of course one of the reasons why Bayesian modeling has become so popular in recent years. But the price you pay for being able to fit more complex random effects models in a Bayesian framework is computation time. In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.

@@

In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data set used in Tutorial 1a. To keep it short, we only fitted the effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4. Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024). Again, the resulting conditional accuracy model M1i\_ca\_f did not converge and resulted in a singular fit.

@@

The general approach to planning that we adopt here involves simulating data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
The basic structure and code follows the examples outlined by Solomon Kurz in his
`power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Fit a regression model to an existing dataset.
\item
  Use the regression model parameters to simulate one new dataset.
\item
  Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
\item
  Summarise the simulated data to estimate likely power or precision of the research design options.
\end{enumerate}

Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data. However, when each model takes several hours to build, and we may want to simulate 1000s of datasets, it can be computationally demanding for desktop machines. So, for ease, here we just use the raw simulated datasets to guide future expectations.

@@

We again use the data from Panis and Schmidt (2016) to provide a worked example.
We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent). We chose to focus on a subsample of the data to ease the computational burden. We also used a full varying effects structure, with the model formula, as follows:

@@

Now that we have summarised our simulated data, what planning decisions could we make about a future study?
How many trials per condition should we collect and how many participants should we test?
Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022).
There is no straigtforward and clear-cut answer.
Some considerations might be\ldots{}

\begin{itemize}
\tightlist
\item
  How much power or precision are you looking to obtain in this particular study?
\item
  Are you running multiple studies that have some form of replication built in?
\item
  What resources do you have at your disposal, such as time, money and personnel?
\item
  How easy or difficult is it to obtain the specific type of sample?
\end{itemize}

If we were running this kind of study in our lab, what would we do?
We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed in the previously published study that this work is building upon (Panis \& Schmidt, 2016).
Then we might pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80\% power mark.
If we wanted to maximise power based on these simulations, and we had the time and resources available, then we test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.

>>>>

\abstract{%
Time-to-event data such as response times and saccade latencies form a cornerstone of experimental psychology, and have had a widerspread impact on our understanding of human cognition. However, the orthodox method for analysing such data -- comparing means between conditions -- is known to conceal valuable information about the timeline of psychological effects, such as their onset time and duration. The ability to reveal finer-grained, ``temporal states'' of cognitive processes can have important consequences for theory development by qualitatively changing the key inferences that are drawn from psychological data. Luckily, well-established analytical approaches, such as event history analysis (EHA), are able to evaluate the detailed shape of time-to-event distributions, and thus characterise the time course of psychological states. One barrier to wider use of EHA, however, is that the analytical workflow is typically more time-consuming and complex than orthodox approaches. To help achieve broader uptake of EHA, in this paper we outline a set of tutorials that detail one distributional method known as discrete-time EHA. We touch upon several key aspects of the workflow, such as how to process raw data and specify regression models, and we also consider the implications for experimental design, as well as how to manage inter-individual differences. We finish the article by considering the benefits of the approach for understanding psychological states, as well as the limitations and future directions of this work. Finally, the project is written in R and freely available, which means the approach can easily be adapted to other data sets.
}

@@

In this paper, we focus on a distributional method for time-to-event data known as discrete-time Event History Analysis (EHA), a.k.a. survival analysis, hazard analysis, duration analysis, failure-time analysis, and transition analysis (Singer \& Willett, 2003). We hope to show the added value of EHA for knowledge and theory building in cognitive psychology and related areas of research, such as cognitive neuroscience. Most importantly, we provide tutorials that provide step-by-step code and instructions in the hope that we can enable others to use EHA in a more routine, efficient and effective manner.

@@

We provide seven different tutorials, which are written in the R programming language and publicly available on our Github page (\url{https://github.com/sven-panis/} Tutorial\_Event\_History\_Analysis), along with all of the other code and material associated with the project. The tutorials provide hands-on, concrete examples of key parts of the analytical process, so that others can apply EHA to their own time-to-event data. Each tutorial is provided as an RMarkdown file, so that others can download and adapt the code to fit their own purposes. Additionally, each tutorial is made available as a .html file, so that it can be viewed by any web browser, and thus available to those that do not use R. Finally, the manuscript itself is written in R using the papaja package (Aust \& Barth, 2024), which makes it computationally reproducible, in terms of the underlying data and figures.

@@

We recommend several excellent textbooks for a comprehensive background context to EHA (Allison, 2010; Singer \& Willett, 2003) and for a more general introduction to understanding regression equations (Gelman, Hill, \& Vehtari, 2020; Winter, 2019). Our focus here is not on providing a detailed account of the underlying regression equations, since this topic has been comprehensively covered many times before. Instead, we want to provide an intuition regarding how EHA works in general, as well as in the context of experimental psychology. As such, we only supply regression equations in section D of the supplementary material.

@@

Figure 3 shows that the early upswing in hazard is equal for both priming conditions (Figure 3, upper panel), and that early emitted responses are always correct in the congruent condition and always incorrect in the incongruent condition (Figure 3, lower panel). These results show that for short waiting times (\textless{} bin 6), responses always follow the prime (and not the target, as instructed). During timebin 6 the target-triggered response channel is activated and causes response competition -- ca(6) = .5 -- and a lower hazard probability in the incongruent condition. For waiting times of 600 ms or more, the hazard of response occurrence is lower in incongruent compared to congruent trials, and all responses emitted in these late bins are correct.

@@

We perform data wrangling following tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
In short, functional programming means that you avoid writing your own loops and instead use functions that have been built and tested by others.
In addition, we also supply a set of custom-built functions, which make the process of data wrangling in the context of data preparation and descriptive statistics a lot quicker and more efficient.

@@

Our lab adopts an estimation approach to multilevel regression (Kruschke \& Liddell, 2018; Winter, 2019), which is heavily influenced by the Bayesian framework as suggested by Richard McElreath (Kurz, 2023b; McElreath, 2020). We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013). This means that wherever possible we include varying intercepts and slopes per participant.
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria (e.g., WAIC) and cross-validation (e.g., LOO), to evaluate out-of-sample predictive accuracy (McElreath, 2020). We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

@@

In contrast, the `person-trial-bin' data (Table 2) has a different, more extended structure, which indicates in which bin a response occurred, if at all, in each trial. Therefore, the `person-trial-bin' data generates a 0 in each bin until an event occurs and then it generates a 1 to signal an event has occurred in that bin. This data set is used when fitting hazard models (Tutorials 2a and 3a). It is worth pointing out that there is no requirement for an event to occur at all (in any bin), as maybe there was no response on that trial or the event occurred after the time window of interest. Likewise, when the event occurs in bin 1 there would only be one row of data for that trial in the person-trial-bin data set.

@@

\caption{Estimated discrete-time hazard (h), survivor (S), conditional accuracy (ca) and probability mass (P) functions for participant 6. Vertical dotted lines indicate the estimated median RTs. Error bars represent +/- 1 standard error of the respective proportion.}\label{fig:eha-plot}
\end{figure}

@@

These distributional results suggest that participant 6 is initially responding to the prime even though (s)he was instructed to only respond to the target, that response competition emerges in the incongruent prime condition around 300 ms, and that only slower responses are fully controlled by the target stimulus. Qualitatively similar results were obtained for the other five participants. When participants show qualitatively similar distributional patterns, one might consider aggregating their data and plotting the group-average distribution per condition (see Tutorial\_1a.Rmd).

@@

In this third tutorial, we illustrate how to fit Bayesian multilevel regression models to the RT data of the masked response priming data used in Tutorial 1a. Fitting (Bayesian or non-Bayesian) regression models to time-to-event data is important when you want to study how the shape of the hazard function depends on various predictors (Singer \& Willett, 2003).

@@

There are several analytic decisions one has to make when fitting a discrete-time hazard model. First, one has to select an analysis time window, i.e., a contiguous set of bins for which there is enough data for each participant. Second, given that the dependent variable (event occurrence) is binary, one has to select a link function (see section C in the supplementary material). The cloglog link is preferred over the logit link when events can occur in principle at any time point within a bin, which is the case for RT data (Singer \& Willett, 2003). Third, one has to choose whether to treat TIME (i.e., the time bin index t) as a categorical or continuous predictor. And when you treat a variable as a categorical predictor, you can choose between reference coding and index coding. With reference coding, one defines the variable as a factor and selects one of the k categories as the reference level. Brm() will then construct k-1 indicator variables (see model M1d in Tutorial\_2a.Rmd for an example). With index coding, one constructs an index variable that contains integers that correspond to different categories (see models M0i and M1i below). As explained by McElreath (2020), the advantage of index coding is that the same prior can be assigned to each level of the index variable, so that each category has the same prior uncertainty.

@@

If we want to estimate the population-level effect of prime type on hazard, we can base our conclusion on Figure 7A. The contrast ``congruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.02, 0.17{]}), and 0.14 hazard units in bin (240,280{]}) (95\% CrI = {[}0.04, 0.25{]}). For the other bins, the 95\% credible interval contained zero.
The contrast ``incongruent minus blank'' was estimated to be 0.09 hazard units in bin (200,240{]} (95\% CrI = {[}0.01, 0.21{]}), -0.19 hazard units in bin (320,360{]} (95\% CrI = {[}-0.31, -0.06{]}), -0.27 hazard units in bin (360,400{]} (95\% CrI = {[}-0.45, -0.04{]}), and -0.23 hazard units in bin (400,440{]} (95\% CrI = {[}-0.40, -0.03{]}). For the other bins, the 95\% credible interval contained zero.

@@

In this fourth tutorial, we illustrate how to fit a Bayesian multilevel regression model to the timed accuracy data from the masked response priming data used in Tutorial 1a. The general process is similar to Tutorial 2a, except that (a) we use the person-trial data, (b) we use the logit link function, and (c) we change the priors. To keep the tutorial short, we only fit one conditional accuracy model, which was based on model M1i from Tutorial 2a and labelled M1i\_ca.

@@

In this fifth tutorial we illustrate how to fit a multilevel regression model to RT data in the frequentist framework, for the data used in Tutorial 1a. The general process is similar to that in Tutorial 2a, except that there are no priors to set.

Again, to keep the tutorial concise, we only fit model M1i (see Tutorial 2a) using the function glmer() from the R package lme4. Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024). The resulting hazard model is called M1i\_f with the appended ``\_f'' denoting a frequentist model.

In Figure 11 we compare the parameter estimates from the Bayesian regression model M1i with those from the frequentist model M1i\_f.

@@

Figure 11 confirms that the parameter estimates from both Bayesian and frequentist models are pretty similar, which makes sense given the close similarity in model structure. However, model M1i\_f did not converge and resulted in a singular fit. This is of course one of the reasons why Bayesian modeling has become so popular in recent years. But the price you pay for being able to fit models with more complex varying effects structures via a Bayesian framework is increased computation time. In other words, as we have noted throughout, some of the Bayesian models in Tutorials 2a took several hours to build.

@@

In this sixth tutorial we illustrate how to fit a multilevel regression model to the timed accuracy data in the frequentist framework, for the data used in Tutorial 1a. To be concise, we only fit effects from model M1i\_ca (see Tutorial 2b) using the function glmer() from the R package lme4. Alternatively, one could also use the function glmmPQL() from the R package MASS (Ripley et al., 2024). The resulting conditional accuracy model, which we labelled M1i\_ca\_f, did not converge and resulted in a singular fit. Again, this just highlights some of the difficulties in fitting reasonably complex varying/random effects structures in frequentist workflows.

@@

The general approach to planning that we adopt here involves simulating reasonably structured data to help guide what you might be able to expect from your data once you collect it (Gelman, Vehtari, et al., 2020).
The basic structure and code follows the examples outlined by Solomon Kurz in his `power' blog posts (\url{https://solomonkurz.netlify.app/blog/bayesian-power-analysis-part-i/}) and Lisa Debruine's R package faux\{\} (\url{https://debruine.github.io/faux/}) as well as the related paper (DeBruine \& Barr, 2021).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Fit a regression model to existing data.
\item
  Use the regression model parameters to simulate new data.
\item
  Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
\item
  Summarise the simulated data to estimate likely power or precision of the research design options.
\end{enumerate}

Ideally, in the above workflow, we would also fit a model to each dataset and summarise the model output, rather than the raw data. However, when each model takes several hours to build, and we may want to simulate many 1000s of datasets, it can be computationally demanding for desktop machines. So, for ease, here we just use the raw simulated datasets to guide future expectations.

@@

We again use the data from Panis and Schmidt (2016) to provide a worked example.
We fit an index coding model on a subset of timebins (six timebins in total) and for two prime conditions (congruent and incongruent). We chose to focus on a subsample of the data to ease the computational burden. We also used a full varying effects structure, with the model formula as follows:

@@

Now that we have summarised our simulated data, what planning decisions could we make about a future study?
More concretely, how many trials per condition should we collect and how many participants should we test?
Like almost always when planning future studies, the answer depends on your objectives, as well as the available resources (Lakens, 2022).
There is no straigtforward and clear-cut answer.
Some considerations might be as follows:

\begin{itemize}
\tightlist
\item
  How much power or precision are you looking to obtain in this particular study?
\item
  Are you running multiple studies that have some form of replication built in?
\item
  What level of resources do you have at your disposal, such as time, money and personnel?
\item
  How easy or difficult is it to obtain the specific type of sample?
\end{itemize}

If we were running this kind of study in our lab, what would we do?
We might pick a hazard ratio of 0.4 or 0.5 as a target effect size since this is much smaller than that observed previously (Panis \& Schmidt, 2016).
Then we might pick the corresponding combination of trial count per condition (e.g., 200) and participant sample size (e.g., N=10 or N=15) that takes you over the 80\% power mark.
If we wanted to maximise power based on these simulations, and we had the time and resources available, then we would test N=20 participants, which would provide \textgreater90\% power for an effect size of 0.5.
